---
title: "decay explore"
author: "Fritz Ebner"
date: "June 8, 2015"
output: word_document
---

we have 2 different decaying components and we want to look at how they interact.  
First look at averaging 2 different time constants and see what the interaction looks like.
we see that simple mean of exponent will yield a plot(blue) that is in between the constituent plots.   try with similar constants, then when one is much smaller than the other.

```{r}
library(permute)
library(ggplot2)
library(reshape2)
set.seed(42)
getdf <- function(k1,k2,b1,b2) {
    x <- 0:100
    k3 <- (k1+k2)/2
    e1 <- b1*(1-exp(-k1*x))
    e2 <- b2*(1-exp(-k2*x))
    #e3 <- (b1+b2)*(1-exp(-k3*x))
    e3 <- e1 + e2
    df <- data.frame(x,e3)
    mod <- nls(e3 ~ b*(1-exp(-k * x)), data = df, start = list(k = 1, b = 1))
    #mod <- nls(e3 ~ (b1+b2)*(1-exp(-k * x)), data = df, start = list(k = 1))
    eres <- coef(mod)[2]*(1-exp(-coef(mod)[1]*x))
    #eres <- (b1+b2)*(1-exp(-coef(mod)[1]*x))
    print(paste("k1",k1," k2",k2," kcombo",coef(mod)[1]))
    print(paste("arith mean", (k1+k2)/2, " geom mean",sqrt(k1*k2)))
    print(paste("weighted geom mean ",exp((b1*log(k1) + b2*log(k2))/(b1+b2))))
    df2 <- data.frame(x,e1,e2,e3,eres)
}
ggplot(df <- getdf(.025,.1,.25,1), aes(df$x)) +                    
    geom_line(aes(y=df$e1), colour="red") +  
    geom_line(aes(y=df$e2), colour="green") +
    geom_line(aes(y=df$e3), colour="blue") +
    geom_line(aes(y=df$eres), colour="black")
```

OK.  lets extend this to a larger sample and see if it generalizes.  
  Sample k from lo to hi and b from lo to hi, within ranges, then randomize ks and bs,
  so that the pair is shuffled.  First plot the set of 100 curves.
```{r}
savKArith <- NULL
savKGeom <- NULL
savKHarm <- NULL
savKs <- NULL
savBs <- NULL
mksamps <- function(blo, bhi, klo, khi, samps) {
    #samps = 10
    #bs<-seq(blo,bhi,length.out=samps)
    #ks<-seq(klo,khi,length.out=samps)
    # just shuffle the parameter lists independently
    #ks <- ks[shuffle(samps)]
    #bs <- bs[shuffle(samps)]
    ks <- runif(samps, klo,khi)
    bs <- runif(samps, blo,bhi)
    savKs <<- ks
    savBs <<- bs
    # mean of Ks
    print(paste("arith mean ks ", savKArith<<-mean(ks)))
    # geom mean of Ks
    print(paste("geom mean ks ",savKGeom<<-exp(mean(log(ks)))))
    # harmonic mean of Ks
    print(paste("harmonic mean ks ",savKHarm<<-1/mean(1/ks)))
    # sum of Bs
    print(paste("sum of bs ",sum(bs)))
    
    thenms <- c("x",paste0("k",format(ks,digits=1),"b",format(bs,digits=2)))
    x <- 0:100
    # now combine ks and bs into pairs of parameters
    df <- data.frame(x)
    for (i in 1:samps){
        tmp <- bs[[i]]*(1-exp(-ks[[i]]*x))
        df <- cbind(df,tmp)
    }
    colnames(df) <- thenms
    df
}
# plot on same grid, each series colored differently -- 
# good if the series have same scale
num=26
blo=165
bhi = 496
klo=0.14
khi=0.64
df <- mksamps(blo,bhi,klo,khi,num)
dfplot <- melt(df ,  id = 'x', variable_name = 'series')
ggplot(dfplot, aes(x,value)) + geom_line(aes(colour = variable))
```
  
  Now sum the resultant curves to get the sum of exponents.  We know this is technically not an exponential, but perhaps there is a close approximation.

```{r}
sums <- rowSums(df[2:num+1])
df <- cbind(df,sums)
dfplot <- melt(df ,  id = 'x', variable_name = 'series')
ggplot(dfplot, aes(x,value)) + geom_line(aes(colour = variable))
```

  ANd now fit with exponential ... see how it looks.  
  
```{r}
mod2 <- nls(sums ~ b*(1-exp(-k * x)), data = df, start = list(k = 1, b = num*bhi))
eres <- coef(mod2)[2]*(1-exp(-coef(mod2)[1]*df$x))
df <- cbind(df,eres)
dfplot <- melt(df ,  id = 'x', variable_name = 'series')
ggplot(dfplot, aes(x,value)) + geom_line(aes(colour = variable))
# fitted b and K are
print(paste("fitK ",coef(mod2)[1], "  fitb ",coef(mod2)[2]))
# vs original arith and geom and harm
print(savKArith)
print(savKGeom)
print(savKHarm)
# Now lets calculate the weigthed geom mean of the Ks.  Weights are Bs
weightedGeom <- exp(sum(savBs*log(savKs))/sum(savBs))
print(weightedGeom)
```